---
title: "Thematic Mapping of Demographic & Socioeconomic Data"
author: "Jason Jones"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  pagedown::html_paged:
    css: ["style/jason-default-fonts.css", "style/jason-default-page.css", "style/jason-default.css"]
    toc: true
    # change to true for a self-contained document, but it'll be a litte slower for Pandoc to render
    self_contained: false
    front_cover: "img/report_cover.png"
chapter_name: "Section:\\ "
# uncomment this line to produce HTML and PDF in RStudio:
#knit: pagedown::chrome_print
---

<script>
class RepeatingTableHeaders extends Paged.Handler {
  constructor(chunker, polisher, caller) {
    super(chunker, polisher, caller);
  }

  afterPageLayout(pageElement, page, breakToken, chunker) {
    // Find all split table elements
    let tables = pageElement.querySelectorAll("table[data-split-from]");

    tables.forEach((table) => {
      // Get the reference UUID of the node
      let ref = table.dataset.ref;
      // Find the node in the original source
      let sourceTable = chunker.source.querySelector("[data-ref='" + ref + "']");
      // Find if there is a header
      let header = sourceTable.querySelector("thead");
      if (header) {
        // Clone the header element
        let clonedHeader = header.cloneNode(true);
        // Insert the header at the start of the split table
        table.insertBefore(clonedHeader, table.firstChild);
      }
    });

  }
}

Paged.registerHandlers(RepeatingTableHeaders);
</script>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dpi = 320,
                      out.width = "100%", fig.align = "center")

# Load packages
library(tidyverse)
library(sf)

# Census geographies
tracts_sf <- read_rds("data/geographies/tracts_sf.rds")

# Load data
pci_county <- read_rds("data/per_capita_income/pci_county.rds")
pci_places <- read_rds("data/per_capita_income/pci_places.rds")

pci_sf <- read_rds("data/per_capita_income/pci_sf.rds")
mfi_sf <- read_rds("data/median_family_income/mfi_sf.rds")
mhv_sf <- read_rds("data/median_home_value/mhv_sf.rds")
bach_sf <- read_rds("data/educational_attainment/ed_attain_tracts_cleaned.rds") %>%
  filter(label == "Bachelor's degree") %>%
  left_join(tracts_sf) %>%
  st_as_sf()

# Load my theme for plots
jason_theme <- theme(
  text = element_text(family = "Roboto"),
  panel.background = element_blank(),
  panel.grid.minor = element_blank(),
  panel.grid.major.y = element_line(color = "grey"),
  panel.grid.major.x = element_blank(),
  axis.ticks = element_line(color = "grey"),
  axis.title = element_blank(),
  plot.caption = element_text(hjust = 1)
  )

chart_colors <- colorRampPalette(c("#112E51", "#97BCE9", "#C1D7F2", "#FFBEA9", "#853A22"))
```

# Per Capita Income {.chapter}

## Variable Definition

## Thematic Maps

```{r}
pci_sf %>%
  ggplot() +
  geom_sf(aes(fill = estimate), color = "white", size = 0.2) +
  scale_fill_gradientn("Estimate", colors = chart_colors(4), na.value = "white",
                       labels = scales::dollar_format()) +
  theme(legend.position = "left") +
  labs(caption = "Source: United States Census Bureau, American Community Survey (ACS)\nAuthor: Jason Jones",
       title = "Per Capita Income",
       subtitle = "Guilford County: Census Tracts") +
  jason_theme +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank())
```

# Median Family Income {.chapter}

## Variable Definition

## Thematic Maps

```{r}
mfi_sf %>%
  ggplot() +
  geom_sf(aes(fill = estimate), color = "white", size = 0.2) +
  scale_fill_gradientn("Estimate", colors = chart_colors(4), na.value = "white",
                       labels = scales::dollar_format()) +
  theme(legend.position = "left") +
  labs(caption = "Source: United States Census Bureau, American Community Survey (ACS)\nAuthor: Jason Jones",
       title = "Median Family Income",
       subtitle = "Guilford County: Census Tracts") +
  jason_theme +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank())
```

# Median Home Value {.chapter}

## Variable Definition

## Thematic Maps

```{r}
mhv_sf %>%
  ggplot() +
  geom_sf(aes(fill = estimate), color = "white", size = 0.2) +
  scale_fill_gradientn("Estimate", colors = chart_colors(4), na.value = "white",
                       labels = scales::dollar_format()) +
  theme(legend.position = "left") +
  labs(caption = "Source: United States Census Bureau, American Community Survey (ACS)\nAuthor: Jason Jones",
       title = "Median Value (Owner-Occupied Housing Units)",
       subtitle = "Guilford County: Census Tracts") +
  jason_theme +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank())
```

# Educational Attainment {.chapter}

## Variable Definition

## Thematic Maps

```{r}
bach_sf %>%
  ggplot() +
  geom_sf(aes(fill = pct), color = "white", size = 0.2) +
  scale_fill_gradientn("Estimate", colors = chart_colors(4), na.value = "white",
                       labels = scales::percent_format()) +
  theme(legend.position = "left") +
  labs(caption = "Source: United States Census Bureau, American Community Survey (ACS)\nAuthor: Jason Jones",
       title = "Bachelor's Degree Attainment: Population 25 Years and Over",
       subtitle = "Guilford County: Census Tracts") +
  jason_theme +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank())
```


# Notes on Methodology {.chapter}

## Divergent Color Palette
For my thematic maps, I selected a diverging color palette to to bring attention to stark differences in low, medium, and high values. I am following best practice recommendations from The United States Census Bureau and have selected the applied colors based on their recommended diverging color palette. I am also trying to account for Web Content Accessibility Guidelines (WCAG) by meeting or exceeding the recommended color contrast ratio minimum. I believe the colors recommended by The Census Bureau meet these guidelines while also striking an acceptable balance with visual appeal. You can find more information online at The Census Bureau's Data Visualization Standards repository (https://xdgov.github.io/data-design-standards/).

